XOR-Some Prolific Pilot 1

### 2021-03-26 analysis/xor-some-prolific-pilot1.Rmd
prettify, add timeSpent and by-item plots

N=10 participants were recruited for this pilot and compensated 2 pounds/participant.

```{r libraries}
library(tidyverse)
```

@@

```{r data}
d <- read_csv("../data/pilots/results_58_xor-some-Prolific-pilot1_N10.csv")
```

@@

Next, we would exclude participants based on their ratings in the main / example trials (and possibly the bot check) according to preregistered exclusion criteria. For now, all the data is considered.
```{r clean}
d_native %>% group_by(submission_id) %>% count(response) %>% View()
d_main <- d_native %>% filter(trial_name != "example")
d_exmpl <- d_native %>% filter(trial_name == "example")
d_critical <- d_main %>% filter(condition == "critical")

@@

Plot test questions by type
```{r}
d_test <- d_main %>% rowwise() %>% filter(condition == "test") %>% 
  mutate(test_condition = substr(test_question, 6, 9),
         test_condition = ifelse(test_condition == "fals", "false", 
                                 ifelse(test_condition == "unce", "uncertain",
                                        test_condition)))

d_test %>% 
  ggplot(., aes(x = test_condition, y = response)) +
  geom_point(size = 2, alpha = 0.3, position = position_jitter(width = 0.1))
# add mean / CI
```

Plot main rel / comp / pri questions by main condition

```{r}
d_critical <- d_critical %>% 
  pivot_longer(c(competence, relevance, prior), 
               names_to = "class_condition", 
               values_to = "prior_class")
d_critical %>% 
  filter(block != "xor" & block != "some") %>%
  ggplot(., aes(x = as.factor(prior_class), y = response)) +
  geom_point(size = 2, alpha = 0.3, position = position_jitter(width = 0.1)) +
    facet_wrap(main_type~class_condition)
```
Plot main rel / comp questions by main condition, separated into with / without critical utterance
```{r}
# add info to already record this
d_critical <- d_critical %>% 
  mutate(w_utterance = ifelse(is.na(critical_question), F, T))

d_critical %>% 
  filter(block != "xor" & block != "some") %>%
  ggplot(., aes(x = as.factor(prior_class), y = response, color = w_utterance)) +
  geom_point(size = 2, alpha = 0.3, position = position_jitter(width = 0.1)) +
    facet_wrap(main_type~class_condition) # get ratings from the respective trials only 
# use block for getting the actual correct ratings, but class_condition for subsetting the prior classification fo the respective predictor 
```

plot inference rating as a function of respective rating of the explanatory factor (think about a plot where one would see an interaction)

>>>>

N=10 participants were recruited for this pilot and compensated 2 pounds/participant. 8 items (4 some + 4 xor) were manually selected for this pilot to get more data points per item.

```{r libraries}
library(tidyverse)
library(tidyboot)
```

@@

```{r data, results='hide'}
d <- read_csv("../data/pilots/results_58_xor-some-Prolific-pilot1_N10.csv")
```

@@

Check the time participants spent overall on the experiment:
```{r responseTime}
overall_timeSpent <- d_native %>% mutate(timeSpent = round(timeSpent, 2)) %>% distinct(timeSpent)

#  summarize(timeCounts = count(timeSpent) / d_native %>% )
ggplot(data = overall_timeSpent, aes(y=timeSpent, alpha = 0.7)) +
  geom_boxplot() +
  ggtitle("Overall time participants took")
```

Next, we would exclude participants based on their ratings in the main / example trials (and possibly the bot check) according to preregistered exclusion criteria. For now, all the data is considered.
```{r clean}
d_main <- d_native %>% select(-age, -botresponse, -comments, -education, -endTime, 
                              -gender, -languages, -optionLeft, -optionRight, -startDate,
                              -startTime, -timeSpent) %>%
  filter(trial_name != "example")
d_exmpl <- d_native %>% select(-age, -botresponse, -comments, -education, -endTime, 
                              -gender, -languages, -optionLeft, -optionRight, -startDate,
                              -startTime, -timeSpent) %>%
  filter(trial_name == "example")
d_critical <- d_main %>% filter(condition == "critical")

@@

Plot test questions by type
```{r}
d_test <- d_main %>% rowwise() %>% filter(condition == "test") %>% 
  mutate(test_condition = substr(test_question, 6, 9),
         test_condition = ifelse(test_condition == "fals", "false", 
                                 ifelse(test_condition == "unce", "uncertain",
                                        test_condition)))
d_test_ci <- d_test %>% group_by(test_condition) %>% 
  tidyboot_mean(column = response)

d_test %>% 
  ggplot(., aes(x = test_condition, y = response)) +
  geom_point(size = 2, alpha = 0.3, position = position_jitter(width = 0.1)) +
  geom_point(data = d_test_ci, aes(x = test_condition, y = mean), color = "red", 
             size = 4) 
```

```{r, include=FALSE}
# Plot main rel / comp / pri questions by main condition. 
d_critical <- d_critical %>% 
  pivot_longer(c(competence, relevance, prior), 
               names_to = "class_condition", 
               values_to = "prior_class")
d_critical %>% 
  filter(block != "xor" & block != "some") %>%
  ggplot(., aes(x = as.factor(prior_class), y = response, color = as.factor(prior_class) )) +
  geom_point(size = 2, alpha = 0.3, position = position_jitter(width = 0.1)) +
    facet_wrap(main_type~class_condition) +
  xlab("Anticipated categorization (high / low)") +
  theme(legend.position = 'none') 
```
Plot main rel / comp / prior questions by main condition and by prior classification of the item (x-axis), separated into with / without critical utterance (shape, color). Only high-competence items were used for the some trials (only low-competence in the xor trials, respectively). This will be balanced in the main study. 
```{r}
# add info to already record this
d_critical <- d_critical %>% 
  mutate(w_utterance = ifelse(is.na(critical_question), F, T),
         block = ifelse(block == "comp", "competence", 
                        ifelse(block == "rel", "relevance", ifelse(block == "pri", "prior", block) )))
d_critical %>% 
  filter(block != "xor" & block != "some") %>%
  filter(block == class_condition) %>%
  group_by(main_type, class_condition, w_utterance, prior_class) %>% 
  summarize(mean_response = mean(response)) -> d_critical_summary

d_critical %>% 
  filter(block != "xor" & block != "some") %>%
  filter(block == class_condition) %>%
  ggplot(., aes(x = as.factor(prior_class), y = response, shape = w_utterance, color = w_utterance)) +
  geom_point(size = 2, alpha = 0.6, position = position_jitter(width = 0.1)) +
  geom_point(data = d_critical_summary, aes(x = as.factor(prior_class), y = mean_response, shape = w_utterance), 
             color = "red", size = 3) +
  xlab("Anticipated categorization of the items (low / high)") +
  facet_wrap(main_type~class_condition) # get ratings from the respective trials only

```

Plot inference ratings as a function of *anticipated* rating of the explanatory factor, similar to the paper: 
``` {r}
d_critical %>% 
  filter(block == "xor" | block == "some") %>%
  #group_by(block, class_condition, prior_class) %>%
  #summarize(mean_rating = mean(response)) %>%
  ggplot(., aes(x = prior_class, y = response)) +
  geom_point(size = 2, alpha = 0.5) +
  geom_line() +
  facet_wrap(block~class_condition) +
  xlab("Anticipated categorization of the items") +
  ggtitle("Mean inference strength ratings by-predictor")

```

Plot mean ratings (across all 10 participants) for each vignette (with its respective condition indicated in the facet title, in the order relevance/competence/prior), in each condition. For relevance and competence, the color indicates whether it was presented without the utterance or with. 
```{r, fig.width=10, fig.height=7}
bar.width = 0.8
d_critical %>% 
  mutate(title = paste(title, exp_condition, sep = "_")) %>%
  filter(block == class_condition | block == "xor" | block == "some") %>%
  group_by(block, title, w_utterance, main_type) %>%
  summarize(mean_rating = mean(response)) %>% 
  ggplot(., aes(x = block, y = mean_rating, fill = w_utterance)) +
  geom_col(alpha = 0.7, width = bar.width, position = position_dodge(width = bar.width)) +
  #geom_point(size = 2, alpha = 0.5, position = position_jitter(width = 0.1)) +
  facet_wrap(main_type~title, ncol = 4) +
  theme(axis.text.x = element_text(angle = 45)) +
  ggtitle("By-item by-question mean ratings")
```

In conclusion, the technical side of the experiment seems to work without glitches; participants perform as expected on the comprehension questions; the relevance / competence / prior ratings seem to be mostly in line with our anticipated categorizations.

One point to note is that on average participants took slightly longer than the expected 15 minutes to complete the experiment.
